\documentclass[10pt]{article}

%\usepackage{fancyhdr}
 
%\pagestyle{headings}
%\markright{John Smith}

\date{}

\usepackage{amsmath}    % need for subequations
\usepackage{graphicx}   % need for figures
\usepackage{verbatim}   % useful for program listings
\usepackage{color}      % use if color is used in text
\usepackage{subfigure}  % use for side-by-side figures
\usepackage[colorlinks=true,citecolor=blue]{hyperref}   % use for hypertext links
\usepackage{lipsum}
\usepackage{url}

\usepackage[margin=1in]{geometry}

\usepackage{graphicx}
\usepackage{balance}
\usepackage{comment}
\usepackage{amssymb,amsmath}
\usepackage{caption}
\DeclareCaptionType{copyrightbox}
\usepackage{subfigure}
\usepackage{enumerate}
\usepackage{color}
\usepackage{titling}
%\usepackage{subcaption}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\tableref}[1]{Table~\ref{tab:#1}}

\newcommand{\compactimg}{\vspace{-12pt}}

\clubpenalty=10000 
\widowpenalty=10000
\setlength{\parindent}{0cm}



\begin{document}
\pagenumbering{gobble}

% \begin{table}
% \color{blue}
% %\color{Emerald}
% \begin{tabular*}{\textwidth}{l r}
% \large\textbf{TEACHING STATEMENT} & 
% \hfill \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
% \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 
% \large\textbf{NIRANJINI RAJAGOPAL}\\
% \hline
% \end{tabular*}

% \end{table}


Dear Faculty Search Committee members\\
blah\\

From\\
Niranjini Rajagopal\\


I am writing in regard to the available Tenure-Track Assistant Professor Position in the Department of Electrical and Computer Engineering at University of California, Los Angeles. I am currently a PhD candidate in Electrical and Computer Engineering at Carnegie Mellon University. I expect to graduate in May 2019. \\

My research lies in the general area of Cyber-Physical Systems. I work across theory and systems, at the intersection of embedded sensing systems, signal processing and estimation. I worked in the industry prior to joining CMU and my research brings an analytical approach to solving real-world problems. I have an expertise in indoor localization, the focus of my dissertation. My research has resulted in publications in top CPS conferences including IPSN, ICCPS, RTAS, and SenSys, spawned a startup, and has won the Microsoft Indoor Localization competition twice. I strongly believe that my unique research methodology of working across theory and systems, my strong analytical approach as well as my experience in building and deploying systems places me in an excellent postion to tackle challenges faced by emerging CPS. 

%My dissertation focus
My research has been supported by the Carnegie Mellon Presidential Fellowship and the Samsung PhD Fellowship. 

I have been a teaching assistant for undergraduate and graduate courses in signals and communication areas at CMU. I am in the unique position of being  graduate teaching fellow at CMU's center for teaching excellence. I started and led an ECE outreach program for high school students. 






I am applying for an assistant professor position in the department of Electrical and Computer Engineering at University of California, Los Angeles. \\

I work across theory and systems of cyber-physical systems.  My thesis focuses on indoor localization. My research has resulted in ...\\
I bring a unique experience of spanning theory and systems.
My future research agenda is to 

In addition to research, I have a strong interest in teaching and mentoring students.\\

I worked in industry for 3 years, and spent 2 summers in industy. I collaborate with industry during my PhD.\\
I am looking forward to 


\end{document}
