\documentclass[10pt]{article}

%\usepackage{fancyhdr}
 
%\pagestyle{headings}
%\markright{John Smith}

\date{}

\usepackage{amsmath}    % need for subequations
\usepackage{graphicx}   % need for figures
\usepackage{verbatim}   % useful for program listings
\usepackage{color}      % use if color is used in text
\usepackage{subfigure}  % use for side-by-side figures
\usepackage[colorlinks=true,citecolor=blue]{hyperref}   % use for hypertext links
\usepackage{lipsum}
\usepackage{url}

\usepackage[margin=1in]{geometry}

\usepackage{graphicx}
\usepackage{balance}
\usepackage{comment}
\usepackage{amssymb,amsmath}
\usepackage{caption}
\DeclareCaptionType{copyrightbox}
\usepackage{subfigure}
\usepackage{enumerate}
\usepackage{color}
\usepackage{titling}
%\usepackage{subcaption}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\tableref}[1]{Table~\ref{tab:#1}}

\newcommand{\compactimg}{\vspace{-12pt}}

\clubpenalty=10000 
\widowpenalty=10000
\setlength{\parindent}{0cm}



\begin{document}
\pagenumbering{gobble}

\begin{table}
\color{blue}
%\color{Emerald}
\begin{tabular*}{\textwidth}{l r}
\large\textbf{TEACHING STATEMENT} & 
\hfill \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 
\large\textbf{NIRANJINI RAJAGOPAL}\\
\hline
\end{tabular*}

\end{table}


I enjoy teaching and working with students. I believe teachers and mentors play an important role in the personal and professional growth of students. I am looking forward to taking on these roles as a faculty.

\paragraph{Teaching Experience:}
At Carnegie Mellon University (CMU), I was a teaching assistant for signals and systems (undergraduate course) two semesters, wireless networks and applications (undergraduate and graduate course), and a guest lecturer for embedded systems (undergraduate course). I have assisted with ECE labs for high school students, and with science, maths and robotics classes for K-12 students through outreach initiatives. I have mentored masters students on research projects.  

\paragraph{Teaching Plan:}
The courses I am most prepared to teach are signals and systems;  digital signal processing; statistical signal processing; inference, estimation and detection; probability and random processes; digital communication; wireless networks and applications; linear systems. Other courses I can teach are embedded systems and control systems.
Some seminar courses I can teach are - Mobile sensing applications (will include augmented reality); localization and mapping (theory and applications); internet-of-things; cyber-physical-systems; %These courses would include paper readings, labs and a project in addition to covering concepts.\\

%The other courses I can teach given some time to prepare the materials are embedded systems; control systems%; information theory.\\
I plan to have a project component  and possibly labs in all my classes. I would introduce discussions on research papers for my graduate classes.

\paragraph{Teaching Philosophy: }
I want students to develop the skills to think critically about solving real-world engineering problems. 
%For this, first
%Irrespective of the class I teach, I want my goal for students are to develop the skills to think about system level design of real-world engineering problems
%hrough my teaching in ECE, I strive for students to develop the skills to think critically about both fundamental aspects and system level design of real-world engineering problems. 
Towards this, first, they should develop a strong foundation from first principles. Second, they should develop a framework to connect various concepts to each other. This includes connecting theory to practice, connecting concepts across courses, and connecting concepts to other domains. Third, they should develop the ability to apply their skills to new problems, and be able to communicate their approach. Depending on the level of the course and students, I adapt the emphasis across these learning goals.\\

%I give some examples of my methods towards this from my teaching.\\

%Through my teaching in ECE, I strive for students to develop the skills to think critically about both fundamental aspects and system level design of real-world engineering problems. First, they should develop a strong foundation from first principles. Second, they should develop a framework to connect various concepts to each other. This includes connecting theory to practice, connecting concepts across courses, connecting concepts to other domains and to the real-world. Third, they should develop the ability to approach a system-level problem. I give some examples of my methods towards this from my teaching.\\ 

%This includes developing a strong foundation, 

%Towards this, first, they should develop a strong foundation from first principles. 

%Second, they should develop a framework to connect various concepts to each other. This includes connecting theory to practice, connect concepts across courses, connecting concepts to other domains and to the real-world. Third, they should develop the ability to approach a new problem. This includes developing the skills to know how to break down the problem, thinking about trade-offs between different approaches, and connecting this to the fundamental concepts and think about limits and constrains on the problem. Based on the course topic and level, I have tried different methods towards these learning goals.\\

%First, students should develop a strong foundation from first principles.
%and the first mathematically rigorous core ECE course at CMU. In this course, 
%One of the methods I adopt is to give students sufficient practice and provide targeted feedback. %I designed homework problems in the signals and the wireless courses. 
While teaching signals and systems, my goal was for students to develop a strong foundation in the mathematical concepts and an intuition of the concepts. 
I often designed homework problems by breaking them down into smaller problems. Students said this was helpful since they were able to know how to connect the concepts to solve the problem. %Sometimes I provided the solution and asked them to solve the problem. This gave them quick feedback on their technique if they .
I conducted frequent quizzes and gave them feedback, and revised my recitations based on the students understanding. %I gave them written feedback on assignments. %This informed me on what to emphasize in recitations and office hours.  
To help students develop an intuition of the concepts. I tried different approaches. %
%Students to develp an intuition of the mathematical concepts. SEe different appraoches, pr different ways of thinking about the same problem/concept.
%While teaching signals and systems, 
%my goal was for students to develop an intuition and understanding of the mathematical techniques and learn to implement them. 
I found that some while some students understood the concepts easier through illustrations, others found it easier with a purely mathematical approach. I included both approaches in my recitations and office hours. 
When students found equations confusing and wrote them incorrectly, rather than correcting them, I spent time with them one-on-one in my office hours and helped them step-through the process of interpreting what their equations meant intuitively and in the real-world. This approach helped them understand the concepts better. 
%I had a special convolution session where we went into a deep dive the topic.  
I also tried an activity in the classroom where students first solved a convolution problem, then worked in pairs and discussed their appraoch then solved another problem individually. We saw an improvement in the class performance in the second problem. One assessment of these methods was that the midterm exam performance on a difficult concept, that I had attempted many different ways to teach, was significantly better than the performance over any of the past years. \\

% Several students found convolution confusing.
%E 

In order to help students make connections between theory and practice in signals and systems,  %Students are more motivated to learn when they know how the content applies beyond the course. %Being able to connect the content to the real-world motivated the students since it gave them an idea of what they could do with these tools. 
%Previously the course was largely theory-based. The second year I taught the course, 
I introduced applied problems in my recitations and the homework assignments. We invited faculty from different areas to talk about their research. These lectures have continued since then, and every year I give a short lecture in the class about how the course content applies to my research.  I was a TA for this course in 2012 and 2015. In 2017 I got an email from the faculty that over the past few years there was a department-wide change in trend with more students taking upper-level signals courses, due to enjoying the signals and systems course. Four of us who have been a TA for this course over the years were acknowledged to have contributed to the change in trend. \\

When a TA for the wireless networks class, I designed new labs for students to experiment with the physical layer channel characteristics using various wireless platforms. 
For one of the labs, I asked students to design their own experiments, make a hypothesis about what they expect to see, and then perform the experiment and explain any differences between their hypothesis and the real-world experiments. In another lab, students experimented with channel state information. Through these labs, students learnt to connect the concepts in the class to the real-world applications and demonstrated their understanding of how the unpredictable nature of the wireless channel requires sophisticated design from the upper layers of the stack. Two groups of students subsequently defined their course project based on their interest in these initial labs. \\ %where they defined an area related to the class and explored a new problem and implemented a solution. In this way, students applied 

%I designed another lab based on CSI monitoring where students learnt about the short-term variations in the channel and had to write and share their own observations by experimenting. Through these experiments early on in the course, students demonstrated their understanding of how the real-world nature of wireless signals affects system design. 


%I mentored the undergraduate students on their course project, where they identified an area related to the class, framed a problem and dived deeper into it.  read papers, presented their findings and . In the process, students learnt to connect concepts from the class to other applications. I also mentored them on their course project, where they defined a new problem \\

%
%While mentoring students for research and course projects where they approach a system-level problem, one of my goals is for them to understand that there can be several ways to solve the problem. In the wireless networking class, I mentored the undergraduate students on their course projects. I guided them in defining a problem, exploring different approaches, implementing a solution and presenting their results.\\ % appStudents defined a problem and I guided them in  where students applied concepts in the class to a new problem and implemented a solution.
%I spend time 

%Other skills and student motivation also play a role in student learning. 
Student motivation drives what they do. When I have flexibility in working with students individually, I first ask them what their goals are and discuss various options for moving ahead. When I was a TA for the wireless networking class, I met each student individually during the first week of class and had a conversation with them about their background courses, what their goals were for this course as well as their future plans. This understanding guided how I worked with them during the course. When I mentor students for research, I start off by defining concrete problems where they can clearly measure their progress, and gain confidence to try out harder problems. Over time we together define the research problems based on their interest. At some point they transition to working independently and start teaching me. One student who worked with me on localization got interested in robotics, and subsequently she worked an independent project in simultaneous localization and mapping algorithms and eventually joined a robotics startup. She co-authored two research papers. Another masters student I mentored on mobile augmented reality got interested in this area and took courses in vision and graphics, and subsequently joined Apple. He is a co-author on one of our papers and was part of our team that won best demo and a localization competition. I also give students feedback on how they communicate their ideas and approaches through writing, presentation or explaining concepts on a whiteboard.

%Another instance is a Masters student whom I mentored and while working on mobile augmented reality, he got interested in the area and took courses in computer vision and graphics, and subsequently joined Apple. Another masters student I mentored on localization algorithms got interested in the area and took courses in robotics and ultimately joined a robotics group with embedded expertise. \\
 
%worked closely with me on localization and was interested in computer vision. We defined some problems together and he mobile augmented reality, got interested in the area, and then took courses in graphics and computer vision. We together defined 

 %For instance, a PhD student was taking the course for their research and an undergraduate student was taking the course to get a feel of this domain, which was new to them. I shared with with the faculty and With this understanding, I helped them tailor their labs, surveys an projects as well as it helped
 \paragraph{Outreach: } 
 At CMU, I have been active in outreach and volunteering for women in STEM. %to introduce ECE to high school women. 
One example is, I took the lead in starting a new program called mobile labs to introduce ECE to high school girls. I wrote a grant to secure funding for this program and co-led this program for two years with another PhD student. I felt it was important that students develop confidence with their first ECE experience. For this, I made sure the labs were designed such that all students could complete certain tasks, and were asked to reflect on what new skills they learnt. We also added bonus labs to challenge the students who wanted to explore further. We actively sought out female undergraduate students to lead most of the labs. One measurable outcome of the program was that one of the high school girls joined CMU ECE's summer program. \\

I have worked with several educational outreach programs for K-12 in both India and US in various roles - raising funds, teaching, leading programs, creating instructor manuals and volunteering. Most of these are avenues where we introduce a new domain to students in a fun hands-on manner. Through these experiences, I have learnt to adapt the teaching style based to the culture, resources available, and the background knowledge of students. \\

I led a student seminar at CMU for two years to create a space where students could give practice talks and receive feedback. In this role, I started panel-based seminars to discuss topics such as teaching, fellowship, internships and job search for students overall professional growth. 

% I then took the lead on starting a mobile-lab version of this program to take it 
%I started an outreach program where we introduced ECE to high school girls. 
%While mentoring masters students, I start by giving them concrete research problem, where they can clearly measure their progress. I also challenge students with hard problems. In the outreach program, we introduced bonus lab assignments. While working with masters students, at some point I transition them to work on problems where they start working independently and teach me. This makes them more confident to try out new directions.  Finally, students have to communicate what they have learnt, in ways other than implementation. When possible, I also give students feedback on other aspects such as their writing, presentation skills or clarity in explaining concepts on a whiteboard. \\

\paragraph{Learning Science: }To further my growth as a teacher, I actively seek avenues to improve my own understanding of effective teaching practices from pedagogical experts at CMU. I am among a select group (around 8 students across CMU) of graduate teaching fellows (GTF) at the Eberly Center for Teaching Excellence and Educational Innovation. We have regular meetings where we discuss research in pedagogy for our own professional growth and I consult for graduate student instructors by observing their classes and providing objective constructive feedback. Being a part of a team of graduate teaching fellows and interacting with students across various departments in CMU has exposed me to teaching styles and approaches very different from what I am exposed to in engineering. I also attended an HCI course on the design of large scale peer-learning systems to learn about ways to engage students through peer-based learning in large classes. For my future classes, I am eager to try out different approaches for teaching. % and engaging students in the classroom.\\

%and learn about the challenges of the real-world wireless channels. %The students had to design their own experiments, predict the 
%In contrast to Signals and Systems, the Wireless Networks and Applications course was practical in nature.  
%One of my goal for students was to understand the challenges of wireless channels from first principles. I designed new labs for students to experiment with the physical layer signals. Students first had to design their own experiments. 
%and some students opted to experiment with the change in signal with distance, some wanted to test the change in packet reception with crowds, and some wanted to test for variation in data rate across devices and environments. Then students had to describe their hypothesis on the effect of real-world on the signals and sketch out graphs of what they expected to see from the experiments. After conducting the expereiment, they had to think about and explain the differences between their hypothesis and the real-world experiments. 
%Another goal is for them to have the 


% they could improve, sometimes their presentation and communication skills.   
%Other skills are necessary. For instance, confidence to try something new.
%Designed labs that could be completed, reflect. \\

%Often start with giving them something they can complete.\\
%Also provide more challenges.\\
%Research, harder problems where they teach me something\\
%Communication is another skill. One-one. In undergraduate and graduate classes, I will include these.\\
%I ask the students why they are here and what they are looking for.  




%I Second, they should develop a framework to connect various concepts to each other. This includes connecting theory to practice, connect concepts across courses, connecting concepts to other domains and to the real-world.

%Several students gave us feedback  
%and apply them by practice in implementing and applying them in solving problems. % them. %be conversant with the mathematical tools,  on what they mean and apply them. %and 
%Students learnt the mathematical tools by practicing them on several problems  gave students sufficient practice on the mathematical tools, I designed problems where they could apply the same concept to different problems. 
%I desgined the assignmTo give students practice in working with the concepts, I designed assignem 
%I coveregedBased on what students were finding confusion, I By getting frequent feedabck on students understanding, I I addressed any concept that was confusion  and when several students struggled with a concept, I 
%When I taught it for the first time, a challenge was that students did not know what they did not know, in terms of mathematical background. 
%While I taught this class the second time, I conducted a quiz on the first day to quickly assess their understanding of required background. %how 
%I then covered the necessary background material through recitations, homework assignments and one-on-one discussions with students in office hours. Subsequently, I conducted short quizzes regularly on important concepts and gave students immediate feedback. During recitations and homework, I designed the problems by breaking them into smaller parts, each with individual concepts. Students said they was helpful to know which concepts they get stuck at and how they connect together. 
%In the wireless networks class, I designed new labs for students to experiment with the physical layer channel characteristics. Students learnt the principles of physical layer 
%We also introduced guest lectures where faculty presented their research in context of this class (which has continued since then), and I gave a lecture as well about how the theory applies in my research, and continue to give it every year in this course. This helped students connect the course to the broader ECE. I was a TA for this course in 2012 and 2015. In 2017 I got an email from the faculty that over the past few years there was a department-wide change in trend with more students taking signals courses such as DSP, due to enjoying the Signals and Systems course.
%Four of us who have been a TA for this course over the years were acknowledged to have contributed to the change in trend. \\ 

% Second, they should develop a framework to connect various concepts to each other. This includes connecting theory to practice, connect concepts across courses, connecting concepts to other domains and to the real-world.

% Previously the course was largely theory-based. The second year I taught the course, we introduced applied problems, and we invited faculty from different areas to talk about their research. I also gave a presentation about how my research connects to the class. These lectures has continued since then. I was a TA for this course in 2012 and 2015. In 2017 I got an email from the faculty that over the past few years there was a department-wide change in trend with more students taking signals courses such as DSP, due to enjoying the signals and systems course. Four of us who have been a TA for this course over the years were acknowledged to have contributed to the change in trend.\\

% Third, they should develop the ability to approach a new problem. This includes developing the skills to know how to break down the problem, thinking about trade-offs between different approaches, and connecting this to the fundamental concepts and think about limits and constrains on the problem.


%I enjoy teaching and mentoring students. 
%I am passionate about teaching and education. 



% I believe 

%  teaching
% mentoring students. I This is one of my reasons in addition to research that draws me to an academic path. 

% My teaching experience at CMU is from being a teaching assistant twice for signals and systems, an undergraduate level course; the teaching assistant for wireless networks and applications, a mix of undergraduate and graduate level course; guest lecturer in embedded systems course; and through my involvement in various K-12 outreach initiatives.

% A common theme in my teaching 


% Good technology is built on strong foundations, requires an understanding of the connections between theory and practice, and a broader understanding of how to approach a new problem.  Through my interaction with students, I strive to facilitate this learning process.
 % understanding of the connections between the various components. 
 % in students.



% Students learning should be built on strong foundations. % across the courses that I teach. 
% When I was teaching the signals and systems course, the biggest challenge was to address the disparity in the mathematical background and learning styles of students. When I taught this class for the second time, I conducted a quiz in the first class to assess the mathematical preparedness of the students. Subsequently, I covered the necessary background material through working with students one-on-one in office hours, recitations and assessing and providing feedback through homeworks. Where possible, I designed the problems to be self-correcting. During recitations and office hours, I asked students what they found most confusing, and went ove that content. I broke down problems into smaller%Another approach I tried was providing them the final solution and ask them to arrive at the solution. This helped them develop their technique to solve the problem. 
% Consistently, students struggle when the topic of convolution is introduced in the course. While talking to students, they said that they convolution equation is non-intuitive and confusing. 
% My goal was for students to understand the equation and the implications, rather than memorize it. 
% One approach I tried was that I held a special recitation just for this topic and 
% %I tried a few different techniques to tackle this. I introduced a think-pair share activity, where the students first solved a convolution problem, and then discussed their solutions with their neighbors and then solved another problem individually. I 
% %went over what all the incorrect formulae that we came across in their quizzes and homeworks and 
% made them think about and interpret what each of them meant mathematically and practically. In the mid-term exams, only 4 out of the 70 students got the convolution equation wrong, and around 55 students solved the problems without major errors - a vast improvement from previous years. 



% Students should be able to connect theory and practice. I introduced applied problems from areas of communications, image and speech processing and sensor signals. We also introduced guest lectures where faculty presented their research in context of this class (which has continued since then), and I gave a lecture as well about how the theory applies in my research, and continue to give it every year in this course. This helped students connect the course to the broader ECE. I was a TA for this course in 2012 and 2015. In 2017 I got an email from the faculty that over the past few years there was a department-wide change in trend with more students taking signals courses such as DSP, due to enjoying the Signals and Systems course.
% Four of us who have been a TA for this course over the years were acknowledged to have contributed to the change in trend. While guest lecturing and introducing control systems in the embedded systems course, I first ask students 

% Should should be able to learn how to approach a problem. 
% Wireless course, I mentored studns on their coarse project. Define their own problem,
% In office hours, I interacted one-one with students, and would help them arrive at solutions - 
% Recitations - different examples - 

% Localization course:
% Goals: Underlying physics and limits of what is possible. 
% Learn the tools 
% Learn how to approach a problem - design space is broad.
% Learn what components skills they require to solve a problem. 
% For instance, in digital comm course, theoretical limits, signal noise model, to learning that there are many ways to design the system and then understand the design factors that play a role.


% Learn the tools, mathematics - learn the contraints


% While teaching the wireless course, I designed new labs for students to experiment with the physical layer wireless channels and One of my goal for students was to understand the challenges of wireless channels from first principles. I designed new labs for students to experiment with the physical layer signals. 

% I asked students to design their own experiments. Design questions they wanted to answer. Make hypothesis and then exprimentally validate. Able to connect the theoretical to practice - think about how to approach a problem.


%Through my teaching, I strive to achieve three goals for students. The first is for them to develop strong foundations and an understanding from first principles. The second learning goal is for them for understand connectons between various components - including theory and practice, and  the ability to break down a problem into smaller components. A related goal is for them to learn that there can be different ways to solve the same problem.


%My goal for students is 

%I enjoy teaching, and thinking about and iterating on how to better design my teaching.  My goal for students in ECE is for them to develop the ability to think about engineering systems with a foundation based on theory and first principles, and rooted in real-world problems and systems. My goal is for them to be able to break down a problem and to be able to think about various components that are require to solve a problem and how they connect to each other. This also requires them to connect the theory and practice and to connect the content in a course to other courses, and disciplines more broadly. \\

%At CMU, I was a Teaching Assistant for the Signals and Systems course two semesters; I conducted recitations, office hours, designed assignments and exams, and the Wireless Networks and Applications course for a semester; I created new labs, held office hours, designed assignments and exams, guided students on course projects and deep dive survey project.\\

% In the Signals and Systems course, my goal was for student to develop the mathematical foundation and framework for analyzing signals and systems and then apply the theory to solve problems in different domains. 
% %This is the first mathematically intensive core course that undergradautes in CMU ECE take. %As a result, 
% %One of the main challenges is that students have varied mathematical backgrounds prior to this course. 
% When I was a TA for this course for the first time, I learnt that the biggest challenge was to address the disparity in the mathematical background and learning styles of students. When I taught this class for the second time, I conducted a quiz in the first class to assess the mathematical preparedness of the students. Subsequently, I covered the necessary background material through office hours, recitations and assessing and providing feedback through homeworks. I broke down problems into smaller parts so students could realize where they get stuck. Several students gave feedback that this approach was helpful. Where possible, I designed the problems to be self-correcting. For example, I would provide them the final solution and ask them to arrive at the solution. This helped them develop their technique to solve the problem. %During the term of the course, another approach I attempted to make the mathematical parts more accessible was adopting visual representations of signals. For instance while introducing the the concept of periodic signals in two-dimensions, students were asked to infer the differences in the frequency representation between images with horizontal and vertical stripes of varying widths. 
% Consistently, students struggle when the topic of convolution is introduced in the course. %While talking to students, they said that they convolution formula is non-intuitive and confusing. 
% My goal was for students to understand the equation and the implications, rather than memorize it. 
% I tried a few different techniques to tackle this. I introduced a think-pair share activity, where the students first solved a convolution problem, and then discussed their solutions with their neighbors and then solved another problem individually. I went over what all the incorrect formulae that we came across in their quizzes and homeworks and made them think about and interpret what each of them meant mathematically and practically. In the mid-term exams, only 4 out of the 70 students got the convolution equation wrong, and around 55 students solved the problems without major errors - a vast improvement from previous years. To enable students to make connections between the theory and practice, I introduced applied problems from areas of communications, image and speech processing and sensor signals. We also introduced guest lectures where faculty presented their research in context of this class (which has continued since then), and I gave a lecture as well about how the theory applies in my research, and continue to give it every year in this course. This helped students connect the course to the broader ECE. I was a TA for this course in 2012 and 2015. In 2017 I got an email from the faculty that over the past few years there was a department-wide change in trend with more students taking signals courses such as DSP, due to enjoying the Signals and Systems course.
% Four of us who have been a TA for this course over the years were acknowledged to have contributed to the change in trend. \\ 

% In contrast to Signals and Systems, the Wireless Networks and Applications course was practical in nature.  %spanning from design of MAC layer protocols, to network traffic to applications of networks in sensing and localization. 
% One of my goal for students was to understand the challenges of wireless channels from first principles. I designed new labs for students to experiment with the physical layer signals. Students first had to design their own experiments. For instance, some students opted to experiment with the change in signal with distance, some wanted to test the change in packet reception with crowds, and some wanted to test for variation in data rate across devices and environments. Then students had to describe their hypothesis on the effect of real-world on the signals and sketch out graphs of what they expected to see from the experiments. They then performed the experiments, and presented their observations as well as had to think about and explain the differences between their hypothesis and the real-world experiments. I designed another lab based on CSI monitoring where students learnt about the short-term variations in the channel. Through these experiments early on in the course, students learnt about the fundamental challenges of wireless channels.

% %Curiosity and having fun is the first step for a student to get interested in a new domain. However, when students have pre-conceived notions about their abilities to be effective in a new domain, curiosity alone is not enough. 
% Over the years, I have worked with several outreach groups. My goal for students in these avenues here has been to inculcate curiosity and the confidence to venture out into a new domain. At CMU, I co-piloted a Mobile Labs outreach program, where we introduced ECE to high school girls in a school where almost none of the girls went to STEM fields. We introduced students to several different areas of ECE with hands-on labs in energy harvesting, LEDs on micro-controllers, audio processing and programming. Here they approached their learning from an applied perspective, and then began to think about how things worked, and the underlying principles. My goal was to design the sessions such that all students complete the lab and could be challenged with harder problems as well. All students were asked to reflect on what new concepts they learnt. After the first year of the program, one of the students enrolled in CMU ECE's summer school. I also worked on a project to teach physics purely with experiments to middle school students in rural India. Here, my big challenge was to design instructor manuals for instructors who would conduct the labs, %who did not have the skills to design the experiment for making students think critically, and the 
% and the students did not have the component skills required for middle school physics. Another learning experience for me for to teach robotics and science to primary school students in a hands-on manner. Through these varied experiences, I learnt that the teaching methodology has to be adapted with the background of teachers and students, and the available resources and objectives that are feasible. \\

% To further my growth as a teacher, I actively seek avenues to improve my own understanding of effective teaching practices from pedagogical experts at CMU. I have been with the Eberly Center for Teaching Excellence and Educational Innovation in a role of a Graduate Teaching Fellow. Being a part of a team of graduate teaching fellows and interacting with students across various departments in CMU has exposed me to teaching styles and approaches very different from what I am exposed to in engineering.\\

% Good technology is built on strong theoretical foundations, requires application of the theory to practice and an understanding of the connections between the various components. I hope that through my interaction with students, I can facilitate this process in students.

% \section{Teaching Plan}
% most prepared to teach:

% I can teach courses in various areas related to signals, communication and embedded systems. The courses I am most prepared to teach are They include mathematical foundations for elctrical engineering, signals and systems, digital communication, wireless communication, digital signal processing, statistical signal processing, probability and random processes, linear systems, embedded systems, estimation and detection theory, random processes in systems. 

% Cour
% Lab courses, project courses;
% seminars;

% and asked the students to infer the differences in the frequency representation based on visual inspection of the images.
%I realized this the first time I TA-ed this course. When I TA-ed this the second time, 
%I designed a quiz that we gave to the students on the first day of class to assess their mathematical background. After finding the areas where students needed help, I conducted extra classes and designed assessements in the first few weeks that prepared the students for the mathematical rigor of the course. 

%have over the years helped..\\ I created notes for this class in 2012, many of which are in use till date.\\
%Talk about revamping several aspects of the class.\\

 %trend has been changing, 

%My third approach was to include both visual and mathematical approaches for solving these problems, as students learned better with one or the other. 
%Students liked this format and over the semesters, the course has evolved to be more applied, and I continue to go back every year and give a talk about how the theory applies to my research. %more faculty, this helped them understand what upper level courses. Subsequently, for three semesters, I present my research in context of the class. I was able \\

%I like to expriment with and try out different approaches and iterate over time, as well as get student feedback on what is working for them. %I re-designed the recitations for this course 
%Finally, some students learn bettern visually, rather than mathematically. I solved problems both ways, and ask them questions such as duration, amplitude. In the mid-term, only 80 - a vast improvement from previous semesters. 


%I believe that to understand and engineer sensing systems for the real-world, 


%and the skills to build systems in practice, discover the challenges and re-iterate. They should also develop the ability to think about how the theory and practice relate to each other. Another goal for students is to be able to connect the theory and practice and to connect how the content relates to other areas, courses, domains. 

% as they may not take the first step at all.  
%Students also often make early judgements about their abilities to be effective in a new domain based on their first few expereinces in the domain. 

% introducing ECE to high school students, and science and maths to middle and primary school students. 
%  and and primary school instruction, introducing them to physics, robotics, and science. Through all these avenues, students first build something, or exeriment with something new, and then begin to think about the underlying principles and componets. 

%Several K-12 outreach,\\
%Main goal has been to develop curiosity \\
%One methods is to design the exercise in a manner where they complete a task, to break the barrier to entry - or any assumptions that the field is hard\\ Different learning styles, important to design 
%Mobile Labs - introduce them to different aspects of energy - physical based - energy harvesting, LEDs on arduino, sound processing - start with what they are already familiar with. They experiement and try out different aspects. \\
%Outreach with Agastya - goal was to inculcate curiosity, science experiment-based introduction to a new concept. Assignments desgined such as they can complete it\\
%Experiment-based learning - wherever possible - experiements develop cuirosity, make them understand how they relate to real-world and flexibility in learning since they can build on...\\
%C-mites robotics for 3rd graders.\\



%to understand the effect of real-world parameters on the signal and packet reception 
%Mention office hours.\\
%In future, I would like to introduce a project in this course, and frame the theoretical. For instance...
%I am interested 

%In contrast to the SandS course, the Wireless Networks and Applications course was more practical.\\
%I approached this problem by helping students engineer systems 



%My first goal for students is to I strive for students to develop a strong mathematical foundation and the ability to 

%Through my teaching 
%I strive to facilitate the process of developing students 
%I enjoy teaching, the process of thinking through the design of courses and seeing students grow as they learn. 
%My teaching philosophy is to facilitate in students the process of thinking critically about systems in the real-world. These systems can be sensing systems, communication systems right from X to Y. Towards this, my goal for students is to understand the mathematical foundations and be able to think about signals and systems from first principles. My second goal is for students to apply theory to practice by building systems/ working hands-on and be able to reason about the challenges and the different possibilities in solving engineering problem. Depending on the course and students, I take either an approach by getting them interested in the practical problem and application, and then diving deeper into the theory, or an approach where they start with the theory and see how they are applied to various domains.
%My goal for students of lower level classes 

%During my time at CMU, I have twice been a Teaching Assistant (TA) for Signals and Systems, an undergraduate level class, been a TA for Wireless Networks and Applications, a mixed class with udergraduate, Masters and PhD students. In addition, over the years I have been part of several outreach programs teaching primary to high school students.\\ %graduate level 

% At CMU, I 

% have twice been a Teaching Assistant for Signals and Systems, an undergraduate level class. My first goal was for students was to learn the mathematical foundations for analysing signals and systems. and to be able to make connections between the theory and real-world systems. 


% While SS the goal was general theory that was applicable to several domains, the next class I TAed was focused on one domain. 
% In order for students to develop an understading of wireless from first priciples, I introduced and designed labs in the course, early on.\\




% Domain: sensing, signals, systems - understanding real-world systems and building them\\

% What skills do I want students to have?\\
% Goals: For students to think critically, first (1) developing the ability to know what component skills they require (2) understanding fundamentals (3) applying to practice and understanding the challenges 
% \begin{enumerate}
% \item Develop a curiosity to understand how the content they are learning applies to everyday life
% \item Understand from first principles
% \item Apply theory to practice by building systems
% \item Be able to assess engineering trade-offs
% \end{enumerate}

% Either an approach where you start with theory and move towards applied, or start with the application to get them interested and get them thinking about the underlying principles.

% Methods:\\
% Introduce applied problems in assignments, give examples from my research, TA\\
% Convolution - explaining what role it has in my current day research, for instance, duration of system impulse response, amplification, etc.\\
% Designing assignments, problems where the math leads up to the application, for instance, is the system invertible, 
% Think-pair-share - understand what is behind the math\\
% Wireless class - designed new labs - physical layer - they can experiment with and understand the unpredictability of the wireless layer to appreciate the need for designing for reliability at the higher layers\\
% Student projects - \\
% Redesigned recitations - \\
% Expose them to applicatins early on.\\

% Assessments:\\
% Quiz - to check how the class was doing and address common problems early-on\\
% Exams in wireless - applied, weekly homework 

% Big challenge:\\
% Working with students with different prior knowledge\\
% SandS: conducted a quiz at the beginning\\
% Wireless grad and undergrad: exams are different - UG tested lesser on applying.\\

% Several K-12 outreach,\\
% Main goal has been to develop curiosity \\
% One methods is to design the exercise in a manner where they complete a task, to break the barrier to entry - or any assumptions that the field is hard\\ Different learning styles, important to design 
% Mobile Labs - introduce them to different aspects of energy - physical based - energy harvesting, LEDs on arduino, sound processing - start with what they are already familiar with. They experiement and try out different aspects. \\
% Outreach with Agastya - goal was to inculcate curiosity, science experiment-based introduction to a new concept. Assignments desgined such as they can complete it\\
% Experiment-based learning - wherever possible - experiements develop cuirosity, make them understand how they relate to real-world and flexibility in learning since they can build on...\\
% C-mites robotics for 3rd graders.\\

% Courses I can teach at undergraduate level - signals and systems, wireless communication, embedded systems, linear systems, control systems,  DSP\\
% Graduate level - estimation, statistical SP or ADSP, wireless SN, wireless n/w\\
% Special courses - localization, sensing systems (theory + applying + building), preliminary robotics course (state estimation + applied)\\
% Special topics (project based): IoT + AR + robotics + SN\\

% Mentoring students:\\
% Two students for 1.5 years, some for a semester. In mentoring students, I enjoy both, teaching as well as getting them introduced to research. \\Challenge is in balancing giving them independence and guidance. \\
% I believe that each student is unique and it takes one-one mentoring to really help push them and..\\
% Initially, teaching, often this is the first time they are doing a task that \\
% My goal is to push them to do more and eventually they teach me\\

% Through all my teaching I have realized that given the time, good amount of independence for students to be creative, and a good amount of guidance, we can help students..\\





\small
\bibliographystyle{alpha}

\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case

\end{document}
